Post CMOS technologies: An Overview and Analysis

### 2025-06-13 Paper.tex
finished

%% This is file `sample-sigconf-authordraft.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-authordraft.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,authordraft]{acmart}
\usepackage{float}
\usepackage{placeins}
\title{Post CMOS technologies: An Overview and Analysis}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation email}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}

@@

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

@@

Finally, CNTs are among the strongest and stiffest materials 
known, possessing exceptional tensile strength and elastic 
modulus. Moreover, they are remarkably flexible, capable of 
withstanding considerable mechanical strain without fracturing. 
This flexibility is particularly advantageous for emerging 
flexible electronics and 3D integration.

\subsection{Carbon Nanotube Field Effect Transistors (CNFETs)}
The ability of semiconductor CNTs to conduct extremely effectively 
and their nanoscale dimensions make them ideal candidates for 
making the channel in field effect transistors. These properties 
allow CNFETs to offer several significant advantages over silicon 
based FETs:
\begin{itemize}
  \item Aggressive Scaling: Their nanoscale dimensions allow for
  device scaling below the 5 nanometer limit.
  \item High Carrier Mobility: The ballistic or near-ballistic 
  transport of carriers within CNTs leads to very high carrier
  mobility, which translates directly to faster switching speeds 
  and improved device performance.
  \item Lower Power Consumption: CNFETs can operate effectively 
  at lower voltages due to their superior electrical 
  characteristics, resulting in significantly reduced power 
  dissipation.
  \item Enhanced Gate Control: As one-dimensional materials, 
  CNTs offer excellent electrostatic control by the gate, 
  leading to lower off-state currents and steeper subthreshold 
  swings.
  \item Superior Thermal Management: Their high thermal 
  conductivity facilitates efficient heat dissipation, 
  mitigating hot spots and improving reliability in densely 
  packed circuits.
  \item Material Flexibility: The mechanical flexibility of CNTs 
  enables the development of bendable electronics and facilitates 
  potential three-dimensional chip stacking and integration, 
  offering new avenues for compact and innovative designs.
\end{itemize}

\begin{figure}[h]
  \centering
  \includegraphics[width = \linewidth]{images/Screenshot 2025-05-26 at 8.51.02 PM.png}
  \caption{Diagram of CNFET as combination of MOSFET and FinFET}
\end{figure}

@@

On this microprocessor they were able to implement many standard logic 
blocks like multiplexers, arithmetic logic units, decoders, and encoders. 
Most notably, they were able to execute the famous "Hello World"
program.

\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{images/Screenshot 2025-05-26 at 8.56.27 PM.png}
  \caption{Architecture and Design of RV16X-NANO}
\end{figure}

\subsection{Challenges in CNFET Fabrication and Integration}
Despite the considerable amount of progress made, several challenges 
still need to be addressed before CNFETs can achieve widespread 
viability and meet yield requirements for very large scale integration
(VLSI). Most signficantly, material and manufacturing defects are 
the current biggest challenge. Current methods of CNT fabrication 
lack a reliable way of controlling the chirality leading to many
CNTs exhibiting metallic properties instead of being semiconductors. 
Even after achieving semiconducting CNTs, manufacturing processes still
introduce defects and variability across a wafer. This includes variation
in CNFET density and contact resistance leading to non uniform performance
across an entire chip. In addition to manufacturing issues there 
still lacks a way of seamlessly integrating with traditional silicon
CMOS components. Overcoming these challenges is crucial to achieving

@@

\subsubsection{Principle of Operation}
An input logic is provided through a voltage and electric field which 
will set the magnetization in a magnet to be in a certain direction. This 
converts an electrical input into a magnetic state where the spin state 
can now be used to represent the binary 0 or 1.

To then generate an output a current is injected into the nanomagnet 
which will cause the output current from the nanomagnet to all have 
the same

@@

In addition to these new computing regimes they also offer advantages 
in several other areas. They can be manufactured into nanoscale 
dimensions below those of a silicon transistor allowing for greater scaling 
and higher density integration. Their ability to store memory even 
when power is switched enables lower power consumption. Since they can 
also be programmed into continuous resistance values they have the 
ability to store more than just 1 bit per device. Recent experiments
have shown one memristor being able to store 5-7 bits. Advances 
in memristors have also demonstrated a high level of CMOS compatability 
allowing them to be used in hybrid structures with traditional CMOS 
devices.

@@

@@

The current from the differential part then goes to the sensor. 
The sensor compares a combined current coming from either one or 
multiple differential parts and sees if its above the threshold. 
Based on this comparison either a 1 or 0 is outputted indicating 
whether the sum of the weights have surpassed the threshold. 
%\cite here%

\subsubsection{Challenges}
Despite the theoretical efficiency of memristor based TLGs, 
their practical implementation still face significant challenges. 
Fabrication is still very difficult and there are no EDA tools 
compatible with designing devices made of memristors TLGs.

@@

These raw analog values are then fed to a analog to digital converter
(ADC) to then turn into digital magnitudes. Finally, an activation
function (often a sigmoid or ReLU) is applied to these results to 
introduce nonlinearities, making the output resemble various desired 
distributions. 
\begin{figure}[H]
  \centering
  \includegraphics[scale = 0.3]{images/Neural-network-implementation-using-memristor-crossbar-arrays-a-shows-a-conventional.png}
  \caption{Memristor crossbar array}
\end{figure}

\subsubsection{Demonstrations of Memristor Neural Networks}
Notably, the University of Michigan created a memristor based 
computer that utilized a 54x108 memristor crossbar array. 
With this computer they were able to create a single layer perception 
that classified a 5x5 array of pixels, identifying the greek letters 
drawn in the 25 pixel array with 100\% accuracy. In addition, they 
implemented a two layer neural network that found commonalities and 
differences in breast cancer screenings, correctly classifying 
the cases as benign or malignant with 94.6\% accuracy.
%cite here%

@@

\subsection{Personal Thoughts}
With the advent of

\section{Comparative Analysis}

\section{Conclusion}

\section{Acknowledgments}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|\section|, \verb|\subsection|, \verb|\subsubsection|,
\verb|\paragraph|, and \verb|\subparagraph|. The sectioning levels up to
\verb|\subsusection| should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

Below are examples of sectioning commands.

\subsection{Subsection}
\label{sec:subsection}

This is a subsection.

\subsubsection{Subsubsection}
\label{sec:subsubsection}

This is a subsubsection.

\paragraph{Paragraph}

This is a paragraph.

\subparagraph{Subparagraph}

This is a subparagraph.

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
%   \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

>>>>

%% This is file `sample-sigconf-authordraft.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-authordraft.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{acmart}
\usepackage{float}
\usepackage{placeins}
\title{Post CMOS technologies: An Overview and Analysis}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2025}
\acmYear{2025}
%% These commands are for a PROCEEDINGS abstract or paper.
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}

@@

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Post-CMOS Technology, Carbon Nanotubes, CNFETs, Spintronics, MESO Devices, 
Memristors, Crossbar Array, In-Memory Computing, Vector Matrix Multiplication, Majority Gate,
3D Integration, Threshold Logic, Analog Computing}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

@@

Finally, CNTs are among the strongest and stiffest materials 
known, possessing exceptional tensile strength and elastic 
modulus. Moreover, they are remarkably flexible, capable of 
withstanding considerable mechanical strain without fracturing. 
This flexibility is particularly advantageous for emerging 
flexible electronics and 3D integration \cite{appenzeller2008carbon}.

\subsection{Carbon Nanotube Field Effect Transistors (CNFETs)}
The ability of semiconductor CNTs to conduct extremely effectively 
and their nanoscale dimensions make them ideal candidates for 
making the channel in field effect transistors. These properties 
allow CNFETs to offer several significant advantages over silicon 
based FETs:
\begin{itemize}
  \item Aggressive Scaling: Their nanoscale dimensions allow for
  device scaling below the 5 nanometer limit.
  \item High Carrier Mobility: The ballistic or near-ballistic 
  transport of carriers within CNTs leads to very high carrier
  mobility, which translates directly to faster switching speeds 
  and improved device performance.
  \item Lower Power Consumption: CNFETs can operate effectively 
  at lower voltages due to their superior electrical 
  characteristics, resulting in significantly reduced power 
  dissipation.
  \item Enhanced Gate Control: As one-dimensional materials, 
  CNTs offer excellent electrostatic control by the gate, 
  leading to lower off-state currents and steeper subthreshold 
  swings.
  \item Superior Thermal Management: Their high thermal 
  conductivity facilitates efficient heat dissipation, 
  mitigating hot spots and improving reliability in densely 
  packed circuits.
  \item Material Flexibility: The mechanical flexibility of CNTs 
  enables the development of bendable electronics and facilitates 
  potential three-dimensional chip stacking and integration, 
  offering new avenues for compact and innovative designs \cite{javey2003ballistic}.
\end{itemize}

\begin{figure}[h]
  \centering
  \includegraphics[width = \linewidth]{images/Screenshot 2025-05-26 at 8.51.02 PM.png}
  \caption{Diagram of CNFET as combination of MOSFET and FinFET (Image from \cite{appenzeller2008carbon})}
\end{figure}

@@

On this microprocessor they were able to implement many standard logic 
blocks like multiplexers, arithmetic logic units, decoders, and encoders. 
Most notably, they were able to execute the famous "Hello World"
program.
\cite{hills2019modern}

\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{images/Screenshot 2025-05-26 at 8.56.27 PM.png}
  \caption{Architecture and Design of RV16X-NANO (Image from \cite{hills2019modern})}
\end{figure}

\subsection{Challenges in CNFET Fabrication and Integration}
Despite the considerable amount of progress made, several challenges 
still need to be addressed before CNFETs can achieve widespread 
viability and meet yield requirements for very large scale integration
(VLSI). Most signficantly, material and manufacturing defects are 
the current biggest challenge. Current methods of CNT fabrication 
lack a reliable way of controlling the chirality leading to many
CNTs exhibiting metallic properties instead of being semiconductors. 
Even after achieving semiconducting CNTs, manufacturing processes still
introduce defects and variability across a wafer. This includes variation
in CNFET density and contact resistance leading to non uniform performance
across an entire chip. In addition to manufacturing issues there 
still lacks a way of seamlessly integrating with traditional silicon
CMOS components\cite{hills2019modern}.
Overcoming these challenges is crucial to achieving reliable CNFETs
capable of making more advanced computers.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width = \linewidth]{images/41586_2018_770_Fig2_HTML.png}
  \caption{MESO Spin Device Diagram illustrating various operating mechanisms (image from 
  \cite{manipatruni2019scalable})}
\end{figure}

\subsubsection{Principle of Operation}
The input logic is provided as an electrical charge or voltage. 
This input is applied to a magnetoelectric switching node, which 
can change the magnetization direction of the nanomagnet using a 
voltage. This effectively converts the electrical input into a 
stable magnetic state (representing a 0 or 1 based on its up or 
down spin), which represents the logic state. This form of 
"collective state switching" offers superior switching energy, 
lower operating voltage, and enhanced logic density compared to 
charge-only approaches.

The output logic is then generated by supplying a current into 
the device. The magnetic state of the nanomagnet then determines 
sets all the ejected electrons to be strongly polarized due to spin
orbit coupling. This spin is then converted back into a charge 
or voltage driving the subsequent logic gates.

\subsubsection{Logic Implementation and Architecture}
The key feature enabling logic implementation with MESO is due to 
their ability to implement majority gates and inverters. MESO devices
inherently suppport majority logic because of the way multiple 
charge inputs can be summed together to influence the spin state
of a nanomagnet. If the spin of the currents sum to be above a 
certain threshold then that will be enough to change the spin state
of the nanomagnet.

An inverter can then be created by having a nanomagnet affect the 
spin polarization of a wire. If a wire is placed close to a nanomagnet, 
the nanomagnet will act as a sort of a filter. The nanomagnet 
will affect the wire is such a way that electrons with the same spin 
as the magnet will be attracted to the nanomagnet while electrons with 
opposite spin are able to pass through. This then effectively 
outputs a spin state that is opposite to the spin state of 
the nanomagnet creating an inverter.

These devices can then be cascaded together to perform more advanced 
logic operations. While there still hasn't been much 
substantial demonstration of this technology the theory is there 
and complicated logic operations are definitely possible.
\cite{manipatruni2019scalable}

\subsection{Current Challenges in Spintronics}
Despite the significant theoretical advantages of MESO devices, there 
still stands considerable challengs to it's implementation. The primary 
hurdle still lies in the fact that these devices cannot be produced reliably. 
Temperature fluctuations also pose a significant threat to MESO devices
as shifts in temperature can alter the spin states of the devices.

\subsection{Personal Thoughts}
While spintronics has a lot of promising benefits with its low power 
and in memory compute, the possibility of an actual computer being 
made with spintronics seems extremeley slim. The current state 
that devices can be manufactured will be nowhere close to the 
reliability needed for spintronics computers. Even in the case that 
an actual computer can be made, it will definitely not be applicable 
to home personal use due to the sensitivity of the devices and how easily 
the spin could change.

The operating principles behind spintronics inherently seems very
volatile. Much like quantum computing, for it to be viable I think 
extensive error correction algorithms will have to be used that will 
take up a signficant amount of the compute available. The principles 
also make it seem very difficult for the devices to be manufactured at 
a scale comparable to current transistors. I think if spintronics 
logic is to become a thing there may need to be different ways to 
perform logic that is more robust and error tolerant and also perhaps 
a bit more simple.

There is also not much action if at all happening in industry 
related to spintronics. Companies are still unwilling to invest 
in which is a big indicator in how far off from deployment the 
technology still is.

Overall, I don't think spintronics will be very viable in performing 
digital logic operations. It may have big potential in spaces like 
memory, but just for logic I don't see it getting very far. If we don't 
see any big advancements in the next 3-5 years I think the research 
and investment around this technology will fall off and people 
will invest all resources to quantum computing.

@@

In addition to these new computing regimes they also offer advantages 
in several other areas. They can be manufactured into nanoscale 
dimensions below those of a silicon transistor allowing for greater scaling 
and higher density integration. Their ability to store memory even 
when power is switched enables lower power consumption. Since they can 
also be programmed into continuous resistance values they have the 
ability to store more than just 1 bit per device. Recent experiments
have shown one memristor being able to store 5-7 bits \cite{papandroulidakis2019practical}. Advances 
in memristors have also demonstrated a high level of CMOS compatability 
allowing them to be used in hybrid structures with traditional CMOS 
devices.

@@

\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{images/Linearly_separable_red-blue_cropped_.svg.png}
  \caption{Figure illustrating linearly separable function}
\end{figure}

@@

The current from the differential part then goes to the sensor. 
The sensor compares a combined current coming from either one or 
multiple differential parts and sees if its above the threshold. 
Based on this comparison either a 1 or 0 is outputted indicating 
whether the sum of the weights have surpassed the threshold. 
\cite{papandroulidakis2019practical}

\subsubsection{Challenges}
Despite the theoretical efficiency of memristor based TLGs, 
their practical implementation still face significant challenges. 
Fabrication is still very difficult and there are no EDA tools 
compatible with designing devices made of memristors TLGs \cite{papandroulidakis2019practical}.

@@

These raw analog values are then fed to a analog to digital converter
(ADC) to then turn into digital magnitudes. Finally, an activation
function (often a sigmoid or ReLU) is applied to these results to 
introduce nonlinearities, making the output resemble various desired 
distributions \cite{aguirre2024hardware}. 
\begin{figure}[H]
  \centering
  \includegraphics[width = \linewidth]{images/Neural-network-implementation-using-memristor-crossbar-arrays-a-shows-a-conventional.png}
  \caption{Memristor crossbar array (image from \cite{hassan2018real})}
\end{figure}

\subsubsection{Demonstrations of Memristor Neural Networks}
Notably, the University of Michigan created a memristor based 
computer that utilized a 54x108 memristor crossbar array. 
With this computer they were able to create a single layer perception 
that classified a 5x5 array of pixels, identifying the greek letters 
drawn in the 25 pixel array with 100\% accuracy. In addition, they 
implemented a two layer neural network that found commonalities and 
differences in breast cancer screenings, correctly classifying 
the cases as benign or malignant with 94.6\% accuracy.
\cite{cai2019fully}

\begin{figure}[h]
  \centering
  \includegraphics[width = \linewidth]{images/Screenshot 2025-06-07 at 8.22.50 PM.png}
  \caption{Illustration of multiple advances achieved by UMich memristor computer (image from \cite{cai2019fully})}
\end{figure}

@@

\subsection{Personal Thoughts}
With the advent of AI and the need for massive amounts of compute 
I see memristors having a big potential to fill this opportunity 
to create more powerful and efficient asics for machine learning. 
The development in recent years has also been pretty promising with 
UMich demonstrating the memristor computer implementing a 
single layer perceptron.

There also seems to be quite a few startups in the space
indicating that memristors may have gotten to the point where 
industry is willing to invest in them. With all the AI startup 
hype these companies are getting massive amounts of money 
and it will boost these companies and the overall memristor fields 
capability to make a breakthrough.

Outside of making neural networks, I don't think memristors have 
that much viability. The current methods for creating logic gates 
with memristors need quite a lot of traditional CMOS technology to 
support it that it just seems like too much effort for not enough 
benefit. The in memory compute promise also doesn't look too strong
yet and until there can be a way to store significant amounts of memory with 
memristors in the compute block I don't see general purpose memristor 
computers coming anytime soon.

\section{Comparative Analysis}
This section will then dive into my personal opinions on how these 
technologies compare with each other weighing their respective 
weaknesses and strengths, exploring potential synergies, and consider 
the current landscape of compute and where money and effort is 
being directed.

The three technologies highlighted above are all among the most 
researched post CMOS technologies in the past few decades. However, 
some definitely have more promise than others. There can 
only be a few winners, and inevitably some will have to fade away 
while others will enjoy widespread adoption. So what exactly are the 
technologies that stand out and in what fields do they have potential.

First, I will go over which technology I don't think will have 
widespread adoption. That technology would have to be spintronics. 
The main appeal of spintronics is its ability to perform in memory compute. 
This idea is definitely appealing as one of the main bottlenecks that now arise 
scaling compute is the ability to move information to and from memory and compute. 
Spintronics definitely does have its advantanges in this area, however I feel like 
memristors also guarantee this promise in a much more simple and easily conceivable way.

On the logic side, though spintronics does promise more capable logic abilities with 
the potential of representing more than 2 states in a single bit, the implementation 
just seems too complicated to become scalable. Of course it sounds nice to be able 
to implement more states, but if you're not able to scale this than it's not really 
of any use. Even if scalability is achieved the operating principles behind spintronics 
logic like storing spin states in a nanomagnet are too volatile to be used in anything 
outside of specialized supercomputers.

On top of all this, the whole field didn't seem to have too many tangible achievements in 
recent times. Research and theory is great, but if none of it can be turned into an actual 
demonstrable product then it really is just theorizing. Unlike the other two technologies 
that have all produced computers that have been able to accomplish some simple and basic tasks
spintronics has not done any of this.

The technology also doesn't appeal much to the landscape of the industry. With AI being 
all the craze and where all the investment is going, spintroncis doesn't seemed to 
be involved in any of it.

Carbon nanotubes and memristors on the other hand have much more potential while also 
demonstrating significant results already. Their respective strengths in creating 
hyper-efficient transistors and enabling powerful in-memory computing align directly
with the primary challenges and investment trends in the industry today, making them 
the most compelling candidates to lead the post-CMOS era.

Carbon nanotubes in particular are extremely promising. The benefits they can provide 
are a huge advantage of the transistors we already have. Even though they don't 
present a whole new computing paradigm they still amount to a significant improvement 
in the ability to scale compute. Their ballistic transport properties also make them 
very compelling as it presents ways to reduce power scaling, one of the main factors 
limiting our ability to add more transistors on a chip.

The results that have already been realized using CNFETs are nothing short of incredible. 
With the MIT Medical Electronic Device Realization center already building a complete 
16 bit computer the promise of carbon nanotubes is no longer just in theory. It is 
knocking on the doorstep ready to take computing to the next step

Because they are not a completely new computing paradigm that also makes it quite appealing. 
With technologies like spintronics and memristors presenting completely new ways to compute
the industry may become hesitant to switch and it will take a significant amount of effort
for everyone to adopt to something so different. Carbon nanotubes on the other hand can operate
just like normal transistors making the transition much more seamless. The difficulty then just 
lies in the fabrication while the rest of the industry can stay more or less the same.

Perhaps one of the most transformative advantages of carbon nanotubes is their potential 
for true 3D integration. Unlike silicon, which requires high-temperature processing that 
prevents stacking active layers, CNTs can be fabricated at low temperatures. This opens the 
door to building processors vertically, layer by layer, leading to an unprecedented density
of compute and memory interwoven in a single chip. This positions CNTs as the heir apparent 
to silicon for processing tasks.

Finally, memristors who don't have as much potential as carbon nanotubes in the world 
of general have massive potential in the world of specialized AI chips. Their inherent 
ability to do matrix multiplication makes them especially appealing in this day and age 
where training the best AI is what has captivated governments around the world.

Just like carbon nanotubes, the results that have already been shown using memristors have 
been nothing short of incredible. The University of Michigan implmeneting a single layer 
perceptron also demonstrates that memristors are also not just something of theory. If these 
results can be demonstrated on progressively larger scale models then they have a serious 
chance of taking over the AI industry and becoming the dominant way of making chips in
that space.

All of this is without even mentioning their potential for in memory compute. Just like 
spintronics, they have the ability to combine memory and compute in the same module and 
if this can be utilized in addition to their matrix multiplication then memristors also 
have a potential chance of brekaing into the general compute space.

While they are also still quite volatile, just like spintronics, the advancement of 
memristors to become consistent and scalable seem much easier than nanomagnets. Carbon 
nanotubes still stand a step above them in terms of reliability and scalability and are 
ahead in manufacturing maturity.

In the final analysis, the race to succeed CMOS may not produce a single winner, 
but rather a new, specialized hierarchy of technologies. While spintronics appears to 
be a distant prospect due to fundamental hurdles, the paths for carbon nanotubes and 
memristors are becoming increasingly clear. Carbon nanotubes stand out as the most 
pragmatic and powerful successor for general-purpose logic, promising to extend the 
paradigm of Moore's Law through superior efficiency and 3D integration. They represent 
the perfection of the processor as we know it.

However, perfecting the processor alone is not enough. The future of computing, 
particularly in the age of AI, hinges on solving the data bottleneck. This is where 
memristors, with their profound potential for in-memory and analog computing, will 
dominate. There is a possibility of a hybrid future where efficient logic chips built 
from carbon nanotubes work in seamless integration with memristor based accelerators 
and memory modules. Dividing the labor and letting each technology excel at its 
inherents strengths present a logical and promising path forward.

\section{Conclusion}
The relentless march of progress in computing has driven 
silicon-based CMOS technology to its fundamental physical limits, 
necessitating a paradigm shift towards novel materials and 
architectural approaches. This paper has explored three 
cutting-edge technologies—Carbon Nanotubes (CNTs), Spintronics, 
and Memristors—each offering distinct and powerful solutions to 
the challenges of performance, power consumption, and 
architectural bottlenecks inherent in current computing systems.

Carbon Nanotubes stand as a compelling successor to silicon in 
transistor technology. Their nanoscale dimensions, exceptional 
carrier mobility, and superior thermal properties enable the 
creation of CNFETs that promise higher speeds, lower power 
consumption, and greater scalability, as evidenced by the 
development of functional CNFET-based microprocessors. While 
significant fabrication hurdles related to material purity and 
manufacturing uniformity remain, the ongoing advancements 
underscore their potential to extend the capabilities of 
traditional digital logic.

Spintronics, by harnessing the intrinsic spin of electrons, 
introduces a revolutionary approach to logic and memory. 
Magnetoelectric Spin-Orbit (MESO) devices demonstrate the 
ability to perform ultra-low-power logic operations by converting 
charge to magnetic states and back, enabling efficient information 
processing with minimal energy dissipation. The inherent 
capability to implement majority gates and inverters fundamentally 
redefines logic design, paving the way for highly dense and 
energy-efficient circuits. Overcoming challenges in gate 
reliability and material integration will be crucial for 
widespread adoption.
